---
title: "RMS Titanic"
author: "(Your name here)"
date: 2020-
output:
  github_document:
    toc: true
---

*Purpose*: Most datasets have at least a few variables. Part of our task in analyzing a dataset is to understand trends as they vary across these different variables. Unless we're careful and thorough, we can easily miss these patterns. In this challenge you'll analyze a dataset with a small number of categorical variables and try to find differences among the groups.

*Reading*: (Optional) [Wikipedia article](https://en.wikipedia.org/wiki/RMS_Titanic) on the RMS Titanic.

<!-- include-rubric -->
# Grading Rubric
<!-- -------------------------------------------------- -->

Unlike exercises, **challenges will be graded**. The following rubrics define how you will be graded, both on an individual and team basis.

## Individual
<!-- ------------------------- -->

| Category | Unsatisfactory | Satisfactory |
|----------|----------------|--------------|
| Effort | Some task __q__'s left unattempted | All task __q__'s attempted |
| Observed | Did not document observations | Documented observations based on analysis |
| Supported | Some observations not supported by analysis | All observations supported by analysis (table, graph, etc.) |
| Code Styled | Violations of the [style guide](https://style.tidyverse.org/) hinder readability | Code sufficiently close to the [style guide](https://style.tidyverse.org/) |

## Team
<!-- ------------------------- -->

| Category | Unsatisfactory | Satisfactory |
|----------|----------------|--------------|
| Documented | No team contributions to Wiki | Team contributed to Wiki |
| Referenced | No team references in Wiki | At least one reference in Wiki to member report(s) |
| Relevant | References unrelated to assertion, or difficult to find related analysis based on reference text | Reference text clearly points to relevant analysis |

## Due Date
<!-- ------------------------- -->

All the deliverables stated in the rubrics above are due on the day of the class discussion of that exercise. See the [Syllabus](https://docs.google.com/document/d/1jJTh2DH8nVJd2eyMMoyNGroReo0BKcJrz1eONi3rPSc/edit?usp=sharing) for more information.

```{r setup}
library(tidyverse)

df_titanic <- as_tibble(Titanic)
```

*Background*: The RMS Titanic sank on its maiden voyage in 1912; about 67% of its passengers died.

# First Look
<!-- -------------------------------------------------- -->

__q1__ Perform a glimpse of `df_titanic`. What variables are in this dataset?

```{r q1-task}
glimpse(df_titanic)
```

**Observations**:

- This appears to be a set of categories, as opposed to a row for each passenger. For example "1st class male child who died" appears to have an n of 0, whereas "3rd class male child who died" has an n of 35
- There's a dominant group with 670 people, the next biggest is 387 and beyond that they're all less than 200

__q2__ Skim the [Wikipedia article](https://en.wikipedia.org/wiki/RMS_Titanic) on the RMS Titanic, and look for a total count of passengers. Compare against the total computed below. Are there any differences? Are those differences large or small? What might account for those differences?

```{r q2-summary}
## NOTE: No need to edit! We'll cover how to
## do this calculation in a later exercise.
df_titanic %>% summarize(total = sum(n))
```

**Observations**:

- "Titanic's passengers numbered approximately 1,317"
- "Titanic had around 885 crew members on board for her maiden voyage"
- 1,317 + 885 = 2202, which is 1 off from the dataset's value of 2201
- With such a small difference there could be a large number of trivial reasons. Perhaps reports were unclear and wikipedia chose to include while the dataset chose to exclude. Perhaps somebody served as crew and as a passenger at different points. Perhaps there was partially missing information on one person, enough to exclude them from the set. Perhaps someone died before the crash.

__q3__ Create a plot showing the count of passengers who *did* survive, along with aesthetics for `Class` and `Sex`. Document your observations below.

*Note*: There are many ways to do this.

```{r q3-task}
df_titanic %>%
    filter(Survived == "Yes") %>%
    ggplot() +
    geom_col(aes(x = Class, y = n, fill = Sex))
```

**Observations**:

- Approximately the same number of 1st class and crew survived, with a slightly lower number of 3rd class survivors and a much lower (~50%) number of 2nd class survivors.
- A higher number of female 1st and 2nd class passengers survived, with an even split in 3rd class. Many more male crew members survived than female.
- This raises a huge number of questions about proportions and base rates, not just the survival numbers. I suspect that 1st class passengers made up a much smaller proportion of the people than 2nd or 3rd class, implying they survived at a higher rate. I suspect there were also many fewer female crew members in general, so the male-heavy split could just reflect that base rate.
- The fact that ~200 crew survived out of 885 is upsetting, as they were presumably the best trained and prepared of the people on board.

# Deeper Look
<!-- -------------------------------------------------- -->

Raw counts give us a sense of totals, but they are not as useful for understanding differences between groups. This is because the differences we see in counts could be due to either the relative size of the group OR differences in outcomes for those groups. To make comparisons between groups, we should also consider *proportions*.[1]

The following code computes proportions within each `Class, Sex, Age` group.

```{r compute-prop}
## NOTE: No need to edit! We'll cover how to
## do this calculation in a later exercise.
df_prop <-
  df_titanic %>%
  group_by(Class, Sex, Age) %>%
  mutate(
    Total = sum(n),
    Prop = n / Total
  ) %>%
  ungroup()
df_prop
```

__q4__ Replicate your visual from q3, but display `Prop` in place of `n`. Document your observations, and note any new/different observations you make in comparison with q3.

```{r q4-task}
df_prop %>%
    filter(Survived == "Yes") %>%
    ggplot() +
    geom_col(aes(x = Class, y = Prop, fill = Sex), position = "dodge")
```

**Observations**:

- This is not super useful, the children's statistics (1/1 survived in 1st class, 5/5 in 2nd) appear to be overriding the columns. Instead of figuring out how to properly treat the ages as one group, I find what's asked in q5 to be a better representation.

__q5__ Create a plot showing the group-proportion of passengers who *did* survive, along with aesthetics for `Class`, `Sex`, *and* `Age`. Document your observations below.

*Hint*: Don't forget that you can use `facet_grid` to help consider additional variables!

This is just a preparatory graph, to show the underlying demographics. These are how many people who died (top row) and survived (bottom row) in each combination. You can see the low number of 1st/2nd class children make those categories less meaningful.

```{r q5-prep-1}
df_prop %>%
    ggplot() +
    geom_col(aes(x = Class, y = n, fill = Sex)) +
    facet_grid(Survived ~ Age)
```


```{r q5-task}
df_prop %>%
    filter(Survived == "Yes") %>%
    ggplot() +
    geom_col(aes(x = Class, y = Prop, fill = Sex), position = "dodge") +
    facet_wrap(~ Age)
```

**Observations**:

- We can see that women always survived at a higher rate, if the 1st/2nd class children are left out. The 3rd class children have the lowest different, with women surviving at a ~40% higher rate. 2nd class looks like the highest different, surviving at 8-9 times the rate.
- 3rd class and crew are predominantly male, because women survive at higher rates but that doesn't show as clearly in the raw numbers of people surviving.
- 3rd class survival rates are average for adult men but significantly lower for all other 3rd class passengers.

# Notes
<!-- -------------------------------------------------- -->

[1] This is basically the same idea as [Dimensional Analysis](https://en.wikipedia.org/wiki/Dimensional_analysis); computing proportions is akin to non-dimensionalizing a quantity.
