---
title: "Timing Data"
author: "Eric"
date: 2020-08-20
output:
  pdf_document
---

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
```

To make time_modes.csv I took data from 3 full log files on 021 (ABBA, ~31G each)
```{r read-csv, message=FALSE, warning=FALSE}
timing <- read.csv("~/Desktop/time_modes.csv") %>%
  group_by(mode) %>%
  mutate(appearances = n()) %>%
  ungroup %>%
  arrange(-appearances) %>%
  mutate(rank = dense_rank(-appearances))
timing %>% glimpse
```

\pagebreak
Here is what the mixture density looks like, very multi-modal.
```{r mix, message=FALSE, warning=FALSE}
timing %>%
  ggplot +
  geom_density(aes(x = time_s))
```

\pagebreak
Here are modes with over 100 observations, you can see how spread they are as well as how a few overlap. Remember that since densities integrate to 1 the height just means that it's a very narrow density, it doesn't indicate the number of samples.
```{r show-modes, message=FALSE, warning=FALSE}
timing %>%
  filter(appearances > 100) %>%
  ggplot +
  geom_density(aes(x = time_s)) +
  geom_density(aes(x = time_s, color = mode))
```

\pagebreak
Here we can see the shape of the density for the top 12 modes  (note that X and Y scales are both free in order to better view the shape). I would say 9/12 are approximately normal.
```{r show-modes-faceted, message=FALSE, warning=FALSE}
timing %>%
  filter(rank <= 12) %>%
  ggplot +
  geom_density(aes(x = time_s, color = mode)) +
  facet_wrap(~ mode + appearances, scales = "free")
```